---
title: "BERD Workshop"
author: "Zachary F. Fisher"
date: "5/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=90)
```

# Preliminaries

We begin by installing and loading the gimme and multivar packages.

```{r}
#install.packages(c("gimme", "multivar"))
library(gimme)
library(multivar)
```

# Example Data

Our running example for the workshop uses data from Rowland and Wenzel (2020). The data describe:

- 125 undergraduates
- 6 measurements per day for 40 days
- Random assignment to mindfulness group or control
- Seven emotion variables:
  - happy, excited, relaxed, satisfied, angry, anxious, depressed, sad

These data were recently made available as part of paper by Haslbeck, Ryan \& Dablander (2022) at the following github site: https://github.com/jmbh/EmotionTimeSeries

You can download the .RDS data file used in this workshop from the following address:

https://github.com/jmbh/EmotionTimeSeries/tree/master/DataClean/Rowland2020

## Read in the .RDS file.

```{r pressure, echo=FALSE}
raw_data <- readRDS("data_Rowland2020.RDS")
head(raw_data)
```
## Prepare Data for Analysis

First, we Save that data as a list where each list element is one individual's data. 

To reduce the runtime of these examples we'll select 10 individuals from the full dataset. The first 5 being from the treatment group and the second five being from the control group. 

```{r}
subj_to_keep <- c(3,25,26,27,28,1,2,6,7,8)
vars_to_keep <- c(
     "happy", 
     "excited", 
     "relaxed", 
     "satisfied", 
     "angry", 
     "anxious", 
     "depressed", 
     "sad"
)
data <- lapply(subj_to_keep, function(i){ 
     raw_data[raw_data$subj_id == i,vars_to_keep] 
})
```


# GIMME: Individual-Level Only Analysis

```{r, message = FALSE, results='hide'}
fit_indsem <- gimme::indSEM(
     data, 
     out = "01_indSEM", 
     standardize = TRUE
)
```

## Plots

```{r}
plot(fit_indsem, file = "subj1")
```
# GIMME: Aggregate-Level Analysis

```{r, message = FALSE, results='hide'}
fit_aggsem <- gimme::aggSEM(
     data, 
     out = "02_aggSEM", 
     standardize = TRUE
)
```

## Plots

```{r}
plot(fit_aggsem)
```


# Standard GIMME


```{r, message = FALSE, results='hide'}
fit_gimmesem <- gimme::gimmeSEM(
     data, 
     out = "03_gimmeSEM", 
     standardize = TRUE
)
```

## Plots 

### Group-Level Results

```{r}
plot(fit_gimmesem)
```
## Individual-Level Results

```{r}
plot(fit_gimmesem, file = "subj1")
```
## Estimates

```{r}
print(fit_gimmesem, file = "subj1")
```
# GIMME: Subgroup Analysis

```{r, message = FALSE, results='hide'}
fit_sub <- gimme::gimmeSEM(
     data, 
     out = "04_sub", 
     standardize = TRUE, 
     subgroup = TRUE
)
```

# GIMME: Confirmatory Subgroup Analysis

```{r, message = FALSE, results='hide'}
subjname <- paste0("subj",1:10)
subgroup <- c(1,1,1,1,1,2,2,2,2,2) #create the subgroup
confirm_dataframe <- data.frame(subjname, subgroup)  #create the dataframe

fit_subcon <- gimme::gimmeSEM(
     data, 
     out = "05_subcon", 
     standardize = TRUE, 
     subgroup = TRUE, 
     confirm_subgroup = confirm_dataframe
)
```
## Plots

### All

```{r}
plot(fit_subcon)
```

### Subgroups

```{r}
plot(fit_subcon, subgroup = 1)
```
```{r}
plot(fit_subcon, subgroup = 2)
```
## Estimates

```{r}
print(fit_subcon, subgroup = 1)
```

```{r}
print(fit_subcon, subgroup = 1, mean = TRUE)
```


# GIMME: Latent Variable Analysis

```{r, message = FALSE, results='hide'}
lv_model <- '
  neg =~ anxious + angry + sad + depressed
  pos =~ happy + excited + relaxed + satisfied
'

fit_latent <- gimme::gimmeSEM(
  data = data,
  out = "06_latent",
  standardize = TRUE,
  lv_model = lv_model
)

```


# multivar

## Data Processing

Missing data routines for multivar are still under development. For that reason we select the first 76 observations for 10 subjects.

```{r, message = FALSE, results='hide', eval = TRUE}
data_all <- lapply(unique(raw_data$subj_id), function(i){ 
     raw_data[raw_data$subj_id == i,vars_to_keep] 
})
data <- lapply(data_all,function(df){df[complete.cases(df),][1:76,]})[1:10]
```

```{r, message = FALSE, results='hide', eval = TRUE}
model <- multivar::constructModel(data = data)
fit <- multivar::cv.multivar(model)
```

## Plot Common Effects

```{r}
plot_results(fit, plot_type = "common")
```
## Plot Total Effects Subject 1

```{r}
plot_results(fit, plot_type = "total", datasets = 1)
```


# References

Haslbeck, J. M. B., Ryan, O., \& Dablander, F. (2022). Multimodality and Skewness in Emotion Time Series. https://doi.org/10.31234/osf.io/qudr6

Rowland, Z., \& Wenzel, M. (2020). Mindfulness and affect-network density: Does mindfulness facilitate disengagement from affective experiences in daily life?. Mindfulness, 11(5), 1253-1266.
